---
title: "House Sales"
author: "Santi"
date: '2022-06-22'
output: pdf_document
---
Let's load the data and run some basic analysis on it.
```{r Load Data}
house_data <- read.csv('ma_lga_12345.csv')
head(house_data, 10)

# Check for possible duplicates and NA values.
sum(is.na(house_data))
sum(duplicated(house_data))

# Run summary on the data
summary(house_data)
```

We can see there doesn't seem to be anything unusual with the data when we look
at the factors individually. Since we are doing time-series analysis, let's 
convert the $saledate$ to a date format and the $bedrooms$ into a factor.

```{r}
house_data$saledate <- as.Date(house_data$saledate, '%d/%m/%Y')
house_data$bedrooms <- as.factor(house_data$bedrooms)
```

Now let's plot $saledate$ with $MA$ (the sale amount). Notices that the 2 and 3
bedroom housing units have two seemingly distinct lines. As we will soon see, 
this is attributed to the fact that they're the only housing units to have sales
data for houses and units i.e codominium. 

```{r}
{plot(house_data$saledate, house_data$MA, xlab = "Sale Date", ylab = "Sale Amount",
     pch = 19, col = c('red','green','grey','black','blue')[house_data$bedrooms])
legend('topleft', c('1','2','3','4','5'), fill = c('red','green','grey','black','blue'))}
```

With that in mind let's filter the $bedroom$ variable to run time-series on both 
types of housing units. We will do the time-series using fiscal quarters.

```{r, warning = FALSE}
library(dplyr)
one_br <- house_data %>% filter(bedrooms == '1')
ts_one <- ts(one_br[,2], start = c(2007,4), frequency = 4)

two_br_unit <- house_data %>% filter(bedrooms == '2', type == 'unit')
ts_two_unit <- ts(two_br_unit[,2], start = c(2007,2), frequency = 4)

two_br_house <- house_data %>% filter(bedrooms == '2', type == 'house')
ts_two_house <- ts(two_br_house[,2], start = c(2007,3), frequency = 4)

three_br_unit <- house_data %>% filter(bedrooms == '3', type == 'unit')
ts_three_unit <- ts(three_br_unit[,2], start = c(2007,3), frequency = 4)

three_br_house <- house_data %>% filter(bedrooms == '3', type == 'house')
ts_three_house<- ts(three_br_house[,2], start = c(2007,1), frequency = 4)

four_br_house <- house_data %>% filter(bedrooms == '4', type == 'house')
ts_four_house <- ts(four_br_house[,2], start = c(2007,1), frequency = 4)

five_br_house <- house_data %>% filter(bedrooms == '5', type == 'house')
ts_five_house <- ts(five_br_house[,2], start = c(2007,3), frequency = 4)
```

Now that we have filtered and created our time-series object by unit type we can 
run our ARIMA model and create a price forecast. Since we will essentially run
the same lines of code for each  time-series object, let's make a function that:
\begin{itemize}
\item Runs ARIMA on the time-series object for each bedroom, unity type combination.
\item Forecasts $n$ points. 
\item Plots the forecast data.
\end{itemize}

```{r warning = FALSE}
library(forecast)
library(Metrics)

housingForecast <- function(time_series, forecast_points, unit_type, brs)
{
  # Run ARIMA and create summary.
  arima_model <- auto.arima(time_series)
  summary(arima_model)
  
  # Forecast the number of points required
  data_forecast <- forecast(arima_model, forecast_points)
  print(data_forecast)
  
  # Plot the forecast data.
  plot(data_forecast, xlab = 'Year', ylab = 'Amount ($)',
       main = paste('Time series forecast for ',brs,'- bdr',unit_type), 
       # This makes our x-axis consistent and more readble.
       xaxp = c(2007, (2020+forecast_points), ((2020+forecast_points) - 2007) %% 1000))
  return(summary(arima_model))
}
```

So now let's do some forcasting.

# One-Bedroom Unit Forecast
```{r}
housingForecast(ts_one, 10, 'unit', 1)
```
# Two-Bedroom Unit Forecast
```{r}
housingForecast(ts_two_unit, 10, 'unit', 2)
```
# Three-Bedroom Unit Forecast
```{r}
housingForecast(ts_three_unit, 10, 'unit', 3)
```
# Two-Bedroom House Forecast
```{r}
housingForecast(ts_two_house, 10, 'house', 2)
```
# Three-Bedroom House Forecast

```{r}
housingForecast(ts_three_house, 10, 'house', 3)
```
# Four-Bedroom House Forecast

```{r}
housingForecast(ts_four_house, 10, 'house', 4)
```
# Five-Bedroom House Forecast

```{r}
housingForecast(ts_five_house, 10, 'house', 5)
```
